Writing this dissertation (and surviving grad school) would have not been possible without the support of many people. My biggest thanks go to my advisor, Judith Degen, who has been the best mentor imaginable. Over the past three years, she has been a constant source of advice, support, encouragement, knowledge, wisdom, and inspiration. It's been incredibly fun working with her so closely on numerous projects, and I hope many more collaborations will follow!

Chris Manning took a leap of faith when he hired me as a research assistant midway through my master's without any serious experience in NLP research. Since then he taught me the basics of doing research and supported me along every step of my graduate career. Thank you for that (and many more things), Chris!

Chris Potts and Dan Lassiter taught my first two linguistics classes, and I am very grateful that years later both of them were part of my dissertation committee. Also thanks to Mike Frank and Tobi Gerstenberg for being on my defense committee. Thanks to Beth Levin for all the advice passed along while drinking tea in the linguistics kitchen, to Vera Gribanova for helping me get through the final stages of the PhD, and to Dan Jurafsky for being the most enthusiastic and kind chair that a linguistics department could have.

The computational linguistics and cognitive science communities at Stanford have been a great place to get inspired and to get feedback on (often half-baked) ideas. Thanks to all current and former members of the NLP group, and in particular to Abi See, Peng Qi, Kyle Mahowald, Jon Gauthier, Gabor Angeli, Spence Green, Siva Reddy, Sonal Gupta, Jesse Mu, and John Hewitt; to the members of the ALPS lab: Brandon Waldon, Daisy Leigh, Michael Hahn, Elisa Kreiss, Leyla Kursat, Eva Portelance, and Ciyang Qing; to Noah Goodman, Mike Frank, Eve Clark, Herb Clark, Tom Wasow, Michael Henry Tessler, and Robert Hawkins.

The linguistics department kitchen and the computational linguistics lab have been similarly great places for many discussions. Thanks to Prerna Nadathur, Simon Todd, Sara Kessler, Sunwoo Jeong, CJ Brickhouse, Emily Lake, and Zion Mengesha for the company at many Friday socials; and thanks to Rob Voigt, Tim Dozat, Dora Demszky, Ywei Luo, Matt Lamm, and Ignacio Cases for being wonderful office mates.

I couldn't have worked on many projects without wonderful collaborators: Masoud Jasbi, Natalia Silveira, Matthew Loder, Yuxing Chen, Sophie Regan, and Philip Weiss.

Grad school wouldn't have been the same without my cohort (and our joint trip to Tahoe): Reuben-Cohn Gordon, Daisy Leigh, Sabrina Grimberg, Branden Chan, and honorary cohort member Ed King.

I was also fortunate to meet many great linguists outside of Stanford. I had many great discussions that improved this dissertation with Chris Barker, Roger Levy, Michael Franke, Dave Kleinschmidt, Rachel Ryskin, Chigusa Kurumada, Sarah Moss, Ted Gibson, Greg Scontras, Molly Babel,  Emily Morgan, Tal Linzen, Greg Scontras, and Judith Tonhauser. And thanks to the entire Universal Dependencies community, and in particular to Marie-Catherine de Marneffe, Joakim Nivre, Miryam de Lhoneux, Teresa Lynn, and Djam\'e Seddah for many great collaborations.

Occasionally, I also did other things than research. Aaron Horvath and Katey Webber showed me the best bike routes around Stanford, Emily Carian (along with Eva Portelance and Christof Brandtner) proved that alphabetic constraints can make food even better, and Daisy Leigh and Ed King hosted the best parties on two continents. Thanks also to Max Hell and Natalie Johnson, Juan, Natasha, Miguel and Nico Pedroza, Jean Lin, Ece Kaynak, David Zuckerman, Matt Franklin, and Annie and Tyler Atura Bushnell for many joint adventures.

My parents, Johannes and Andrea, early on accepted my stubbornness to do things differently that eventually got me to Stanford. My siblings, Veronika, Magdalena, Karoline, and Valentin, are still among my closest friends and provided great excuses to frequently travel to Florida, Berlin, and Vienna. And my grandparents, Johann and Eva, made sure that I wasn't starving during my first year as a master's student.

Many friends in Vienna (and other places in Europe) made sure that I instantly felt at home whenever I visited the Old World: Ronald Malis, Markus Schmeiduch, Arthur Egger, Mona Wahba, Robert Pilgram, Niko Franjikic, Vivian Thiele-Orberg, Sabine Toifelhart, Pia Yazdanpanah, Gudrun Schweighofer, and Anna Wolf.

Eva Portelance joined me when I moved north to San Francisco two years ago. Together with her dog Dora, she created the most beautiful home I've ever lived in, instigated many formidable binge-watching sessions, got me a foster dog named Archie as a distraction from writing my dissertation, and was always there for me in one of the toughest years of my life.  Also, one day I hope to be half as good a cook as her.

From co-parenting Dark and Stormy to visiting the most beautiful places in the US to hosting pie parties, Christof Brandtner and later also Zenelia Paredes were there for almost all the greatest moments over the past eight years. Christof has a unique ability to create a community wherever he goes and my life in the US would not have been filled with as many great people without him.

Alessandra Peter has been the best travel companion, fellow raccoon enthusiast, and Vanillekipferl baking partner. She's also the best listener and nobody else could make me hike a trail in half of the recommended time or do pushups in a desert in Utah. I can't wait to explore New York together with her.

Finally, Markus Svoboda has been my best friend for almost two decades. Despite us not living in the same country for the past ten years, we're still in contact almost every day. He is an excellent mixer of G\&Ts and the best partner for rants about what the left is doing wrong. But even more importantly, he's been a constant pillar of support and always helps me put things in perspective. I dedicate this dissertation to him.
